### 2016-03-07 paper/icfp16-submission/gallery2.tex
10pt font for now

First, we will examine a simple recursive function @sqsum@ that should
square each element of the input list and then compute the sum of the
result. In the following code snippets we will underline the source span
that \ocaml blames for the type error.
%
\begin{ecode}
let rec sqsum xs = match xs with
  | [] -> 0
  | h::t -> __(sqsum t)__ @ (h * h)
\end{ecode}
%
Unfortunately the student has used the list-append operator |@| instead
of \texttt{+} to compute the sum. \ocaml blames the recursive call
\texttt{sqsum t} on line 3 for being an \texttt{int} instead of a
\texttt{list}.
%
\begin{verbatim}
Error: This expression has type
         int
       but an expression was expected of type
         'a list
\end{verbatim}
%
In contrast, \toolname produces a trace showing how the evaluation of
@sqsum [1]@ eventually gets stuck when it tries to evaluate
\texttt{ 0 }|@|\texttt{ 1}.
%
\begin{center}
  \includegraphics[height=125px]{sqsum.png}
\end{center}
%
Note how the eye is drawn to the entire stuck term as opposed to just
the left sub-term. This emphasizes the \emph{conflict} between @int@ and
@list@ rather than assuming one or the other is correct.

Next we consider a function @sumList@ that just sums the elements of the
input list, but has a more unfortunate error.
\begin{ecode}
let rec sumList xs = match xs with
  | []    -> []
  | y::ys -> y + __sumList ys__
\end{ecode}
%
In this case the student has returned @[]@ in the base case instead of
@0@. Unfortunately, \ocaml deduces (incorrectly) from the base case that
@sumList@ must return a @list@, and then proceeds to blame the
\emph{recursive call} on line 3 for producing a @list@ instead of an
@int@.
%
\begin{verbatim}
Error: This expression has type
         'a list
       but an expression was expected of type
         int
\end{verbatim}
%
\toolname's jump compressed trace shows how the evaluation of
@sumList [1; 2]@ gets stuck when it tries to evaluate @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
This is essentially the same error that \ocaml produces, but the trace
clarifies immediately (via the third step) that the @[]@ is the result
of the recursive call @sumList []@, drawing attention to the incorrect
base case.

@@

let _ = wwhile (__f__, 2)
\end{ecode}
%
Unfortunately, the student has forgotten to apply @f@ at all on line 2,
and just matches it directly against a pair. This faulty definition of
@wwhile@ still typechecks however, and \ocaml thus blames the
\emph{call-site} on line 10.
%
\begin{verbatim}
Error: This expression has type
         int -> int * bool
       but an expression was expected of type
         'a * bool
\end{verbatim}
%
\toolname makes no assumptions about the program and quickly draws the
eye to the true error, the @match@ expression on line 2, and highlights
the conflict in matching a function against a pair pattern.
%
\begin{center}
  \includegraphics[height=150px]{wwhile.png}
\end{center}
%

>>>>

First, we will examine a simple recursive function @sqsum@ that should
square each element of the input list and then compute the sum of the
result. In the following code snippets we will underline the source span
that \ocaml blames for the type error.
%
\begin{ecode}
  let rec sqsum xs = match xs with
    | [] -> 0
    | h::t -> __(sqsum t)__ @ (h * h)
\end{ecode}
%
Unfortunately the student has used the list-append operator |@| instead
of \texttt{+} to compute the sum. \ocaml blames the recursive call
\texttt{sqsum t} on line 3 for being an \texttt{int} instead of a
\texttt{list}.
%
\begin{verbatim}
  This expression has type
    int
  but an expression was expected of type
    'a list
\end{verbatim}
%
In contrast, \toolname produces a trace showing how the evaluation of
@sqsum [1]@ eventually gets stuck when it tries to evaluate
\texttt{ 0 }|@|\texttt{ 1}.
%
\begin{center}
  \includegraphics[height=125px]{sqsum.png}
\end{center}
%
Note how the eye is drawn to the entire stuck term as opposed to just
the left sub-term. This emphasizes the \emph{conflict} between @int@ and
@list@ rather than assuming one or the other is correct.

Next we consider a function @sumList@ that just sums the elements of the
input list, but has a more unfortunate error.
\begin{ecode}
  let rec sumList xs = match xs with
    | []    -> []
    | y::ys -> y + __sumList ys__
\end{ecode}
%
In this case the student has returned @[]@ in the base case instead of
@0@. Unfortunately, \ocaml deduces (incorrectly) from the base case that
@sumList@ must return a @list@, and then proceeds to blame the
\emph{recursive call} on line 3 for producing a @list@ instead of an
@int@.
%
\begin{verbatim}
  This expression has type
    'a list
  but an expression was expected of type
    int
\end{verbatim}
%
\toolname's jump compressed trace shows how the evaluation of
@sumList [1; 2]@ gets stuck when it tries to evaluate @2 + []@.
%
\begin{center}
  \includegraphics[height=125px]{sumlist.png}
\end{center}
%
This is essentially the same error that \ocaml produces, but the trace
clarifies immediately (via the third step) that the @[]@ is the result
of the recursive call @sumList []@, drawing attention to the incorrect
base case.

@@

  let _ = wwhile (__f__, 2)
\end{ecode}
%
Unfortunately, the student has forgotten to apply @f@ at all on line 2,
and just matches it directly against a pair. This faulty definition of
@wwhile@ still typechecks however, and \ocaml thus blames the
\emph{call-site} on line 10.
%
\begin{verbatim}
  This expression has type
    int -> int * bool
  but an expression was expected of type
    'a * bool
\end{verbatim}
%
\toolname makes no assumptions about the program and quickly draws the
eye to the true error, the @match@ expression on line 2, and highlights
the conflict in matching a function against a pair pattern.
%
\begin{center}
  \includegraphics[height=150px]{wwhile.png}
\end{center}
%
